---
name: João Cascalheira
date: "`r format(Sys.time(), '%B %Y')`"
qualifications: Ph.D.
position: "Researcher"
pronouns: he/his
address: "ICArEHB, Universidade do Algarve, Campus de Gambelas, 8005-139 Faro, Portugal"
phone: +351 912 468 699
www: www.joaocascalheira.com
email: "jmcascalheira@ualg.pt"
twitter: joaocascalheira
headcolor: "3f51b5"
github: jmcascalheira
output:
  vitae::awesomecv:
    page_total: true
docname: Curriculum Vitae
---

```{r setup, include=FALSE}

options(encoding = "UTF-8")
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(vitae)
library(tidyverse)
library(scholar)
library(bibliometrix)
library(gridExtra)

peer_artic <- bibliography_entries("CV.bib") %>%
  dplyr::filter(type == "article-journal" & !is.na(DOI)) %>%
  dplyr::arrange(!is.na(issued), desc(issued))

n_artic <- bibliography_entries("CV.bib") %>% 
  filter(!type %in% c("book", "thesis", "speech") & is.na(DOI)) %>%
  dplyr::arrange(!is.na(issued), desc(issued))

total_artic <- bibliography_entries("CV.bib") %>% 
  filter(!type %in% c("book", "thesis", "speech"))

oral <- bibliography_entries("CV.bib") %>% 
  filter(type == "speech" & `title-short` == "Talk") %>% 
  arrange(!is.na(issued), desc(issued))

poster <- bibliography_entries("CV.bib") %>% 
  filter(type == "speech" & `title-short` == "Poster") %>% 
  arrange(!is.na(issued), desc(issued))

```

# Summary

I am an archaeologist interested in prehistoric hunter-gatherer adaptations, specifically in how these are related to processes of climate and environmental change. I focus on the study of lithic technological organization and its correlation with various aspects of ecological dynamics and human behavior. I currently direct fieldwork in several sites within the context of my ERC Consolidator Grant - [*FINISTERRA: Population trajectories and cultural dynamics of the late Neanderthals in Far Western Eurasia*](www.finisterra.icarehb.com). I am also a collaborator in different projects on the Middle and Later Stone Age in Mozambique and Sudan. My publication record counts with 5 books/edited volumes, `r nrow(peer_artic)` publications in peer-review journals, and `r nrow(n_artic)` other publications as first author or in collaboration. I also have `r sum(nrow(oral), nrow(poster))` public presentations in international professional conferences. Currently, I am the Director of [*ICArEHB*](www.icarehb.com), a research center for archaeology and the evolution of human behaviour at [*Universidade do Algarve*](www.ualg.pt).

# Education

```{r education}
tribble(
  ~ degree, ~ uni, ~ dates, ~ details,
  "Agregação in Archaeology", "Universidade do Algarve, Portugal", "2024","",
  "Ph.D. in Archaeology", "Universidade do Algarve, Portugal", "2013", "Thesis: The mediterranean influence on the social networks of the Iberian Solutrean",
  "MA in Archaeology", "Universidade do Algarve, Portugal", "2009", "Final classification: 19 out of 20. Thesis: Solutrean lithic technology from the Vale Boi rockshelter",
  "Licenciatura in Cultural Heritage", "Universidade do Algarve, Portugal", "2007", "Final classification: 16 out of 20"
) %>% 
  detailed_entries(with = degree, what = details, when = dates, where = uni)
```

# Employment History

```{r positions}

tribble(
 ~ role, ~ company, ~ dates, ~ topic,
  "Assistant Researcher", "Universidade do Algarve, Portugal", "2019-Present", "Position funded by Fundação para a Ciência e a Tecnologia") %>%
  detailed_entries(with = role, when = dates, where = company, what = topic)

tribble(
 ~ role, ~ company, ~ dates, ~ topic,
  "Post-Doctoral Researcher", "Universidade do Algarve, Portugal", "2014-2019", "Position funded by Fundação para Ciência e a Tecnologia") %>%
  detailed_entries(with = role, when = dates, where = company, what = topic)

```

# Publications

```{r publications, echo = FALSE, include=FALSE}

scholar <- scholar::get_profile("bc9Ul_cAAAAJ")
history <- scholar::get_citation_history("bc9Ul_cAAAAJ")
pubs <- scholar::get_publications("bc9Ul_cAAAAJ")

# Needs updated file
cascalheira <- convert2df(file = "../CV/wos_30.08.2024.bib", dbsource = "wos", format = "bibtex")

cascalheira <- cascalheira %>% mutate(AU = replace(AU, AU == "MARREIROS J;BICHO N;GIBAJA J;PEREIRA J", "MARREIROS J;BICHO N;GIBAJA J;PEREIRA J;CASCALHEIRA J"))

wos <- Hindex(cascalheira, field = "author", elements = c("CASCALHEIRA J"), sep = ";", years = 20)$H

```

Citation metrics -- August 2024: **`r scholar$total_cites`** citations \| [Google Scholar](https://scholar.google.pt/citations?user=bc9Ul_cAAAAJ) H-index: **`r scholar$h_index`** \| [Web of Science](%22https://publons.com/researcher/1736137/joao-cascalheira/%22) H-index: **`r wos$h_index`**

```{r publicationsgraph, fig.align = 'center', fig.height=2, fig.width=6}

everysecond <- function(x){
  x <- sort(unique(x))
  x[seq(2, length(x), 2)] <- ""
  x
}


history$year <- as.character(history$year)
n_citations <- ggplot(data=history, aes(x=year, y=cites)) +
  geom_bar(stat="identity", fill = "#3f51b5") +
#  geom_smooth(data = history %>%  group_by(year) %>% summarise(count=sum(cites)),
#              aes(x = year, y = count, group=1),
#              method = "lm", se= FALSE, color = "firebrick1", size = 2) +
  labs(title = "", subtitle = "New citations per year") +
  scale_x_discrete(breaks = history$year, labels = everysecond(history$year), limits = history$year) +
  scale_y_continuous(breaks = c(0, 75, 150)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position="none", axis.ticks = element_blank(), text=element_text(size=13)) +
  ylab("") + xlab("")



total_artic <- as.data.frame(
  total_artic
)


pubs <- total_artic %>%
  mutate(issued = as.character(unlist(issued))) %>% 
#  filter(issued != "2023") %>% 
  drop_na(issued) %>% 
  group_by(issued) %>%  
  summarise(N = n()) %>% 
  rename(year = issued)

n_articles <- ggplot(pubs, aes(year, N)) + 
  geom_bar(stat = "identity", fill = "#3f51b5") +
#  geom_smooth(data = pubs,
#              aes(x = year, y = N, group=1),
#              method = "lm", se= FALSE, color = "firebrick1", size = 2) +
  labs(title = "", subtitle = "New publications per year") +
  scale_x_discrete(breaks = pubs$year,labels = everysecond(pubs$year), limits = pubs$year) +
  scale_y_continuous(limits = c(0, 18), breaks = c(0, 6, 12, 18)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), text=element_text(size=13), axis.ticks = element_blank()) +
  xlab("")

grid.arrange(n_articles, n_citations, nrow = 1)

```

## Books and Edited Volumes

```{r books}

bibliography_entries("CV.bib") %>% 
  filter(type == "book") %>% 
  arrange(!is.na(issued), desc(issued))

```

## Articles in Peer-Reviewed Journals

```{r peerreview}

bibliography_entries("CV.bib") %>%
  filter(type == "article-journal" & !is.na(DOI)) %>% 
  arrange(!is.na(issued), desc(issued))

```

## Other Articles

```{r otherpubs}

bibliography_entries("CV.bib") %>% 
  filter(!type %in% c("book","thesis","speech") & is.na(DOI)) %>%
  arrange(!is.na(issued), desc(issued))

```

# Grants, Awards and Fellowships

Total awarded \~2000,000€

```{r grants}

tribble(
  ~ project, ~ amount, ~ year,
  "FINISTERRA - Population trajectories and cultural dynamics of Late Neanderthals in Far Western Eurasia. Funded by European Research Council Consolidator Grant", "1.899,696€", "2022-",
  "Neanderthals and Early Modern Humans in the Escoural Cave (southern Portugal). Funded by Wenner-Gren Foundation", "18,000€", "2020-",
  "The origins and evolution of human cognition and the impact of Southwestern European coastal ecology. Funded by Fundação para a Ciência e Tecnologia", "240,000€", "2018-",
  "Discovering ancient societies in central Portugal: the Muge shellmiddens. Funded by Earthwatch Institute", "51,378€", "2018-",
  "Investigating the earliest Anatomically Modern Human occupation in Western Algarve. Funded by Institute for Field Research", "21,000€", "2018-",
  "Ballistic performance of Upper Paleolithic stone-tipped projectiles: an assessment using computational fluid analysis. Funded by Archaeological Institute of America", "4,467€", "2017-",
  "Portuguese National Science Foundation Post-Doctoral Fellowship", "107,280€", "2014-19",
  "Portuguese National Science Foundation Ph.D. Fellowship", "47,040€", "2010-13",
  "COST Action TD0902 – Submerged Prehistoric Archaeology and Landscapes of the Continental Shelf. Short Term Scientific Mission: The Mediterranean influence in the social networks of the Final Solutrean of Iberia: the cave of Parpalló", "1,200€", "2012",
  "Paleoanthropology Society – Travel grant to the Paleoanthropology Annual Meeting, May, 2010, St. Louis, US", "400€","2010",
  "Universidade do Algarve – Merit grant", "1,350€", "2009",
  'Portuguese National Science Foundation – Investigation grant within the project PTDC/HAH/64185/2006 – UcaRVaT – “The last hunter-gatherers in the Tejo valey – the Muge shellmiddens”', "8,940€", "2008-09"
) %>%
  brief_entries(project, year, amount)

```

# Main Research Projects

## As Principal Investigator and Co-Investigator

```{r projects}

tribble(
  ~ project, ~ year,
  "FINISTERRA - Population trajectories and cultural dynamics of Late Neanderthals in Far Western Eurasia. Funded by European Research Council Consolidator Grant","2022-",
  "Neanderthals and Early Modern Humans in the Escoural Cave (southern Portugal). Funded by the Wenner-Gren Foundation", "2020-",
  "The origins and evolution of human cognition and the impact of Southwestern European coastal ecology. Co-PI with Nuno Bicho. Funded by Fundação para a Ciência e Tecnologia - ALG-01-0145-FEDER-27833", "2018-",
  "Discovering ancient societies in central Portugal: the Muge shellmiddens. Co-PI with Célia Gonçalves, Lino André and Nuno Bicho. Funded by the Earthwatch Institute", "2018-",
  "Investigating the Earliest anatomically modern human occupation in western Algarve. Co-PI with Nuno Bicho. Funded by the Institute for Field Research", "2018-",
  "Ballistic performance of Upper Paleolithic stone-tipped projectiles: an assessment using computational fluid analysis. Funded by the Archaeological Institute of America", "2017-"
) %>%
  brief_entries(project, year)

```

## As Collaborator

```{r ascollaborator}

tribble(
  ~ project, ~ year,
  "Are Neanderthals adapted to heavy masticatory and paramasticatory function?. PI: Ricardo Godinho. Funded by the Fundação para a Ciência e Tecnologia - 022.07737.PTDC", "2023-",
  "The Muge Shellmiddens Project: A new portal to the last hunter-gatherers the Tagus valley, Portugal. PI: Célia Gonçalves. Funded by the Fundação para a Ciência e Tecnologia - ALG-01-0145-FEDER-29680", "2018-",
  "Climate change during the Quaternary in Inhambane, Southwestern Mozambique, and its role in human evolution. PI: Ana Gomes. Funded by the Fundação para a Ciência e Tecnologia - PTDC/HAR-ARQ/28148/2017", "2018-",
  "Middle and Late Stone Archaeology in Machampane Valley, Limpopo River basin, southern Mozambique. PI: Nuno Bicho. Funded by the National Geographic Society", "2017-18",
  "Stone Age Vilankulos: Modern Human Origins Research south of the Rio Save, Mozambique. PI: Jonathan Haws. Funded by Fundação para a Ciência e Tecnologia", "2016-",
  "Middle Stone Age Archaeology survey in the Limpopo and Elephant rivers valleys, southern Mozambique. PI: Nuno Bicho. Funded by the National Geographic Society/Waitt - W-373-15", "2015-16",
  "Middle Stone Age archaeology and the origins of modern humans in Southern Mozambique. PI: Nuno Bicho. Funded by the Fundação para a Ciência e Tecnologia - PTDC/EPH-ARQ/4998/2012", "2013-15",
  "GIS predictive modeling for the discovery of new Mesolithic sites in Central Portugal. PI: Célia Gonçalves
Funded by the Calouste Gulbenkian Foundation", "2013-2015",
  "The last Neanderthals and the emergence of Modern Humans in Southwestern Iberia. PI: Nuno Bicho. Funded by the Fundação para a Ciência e Tecnologia", "2012-14",
  "On the edge: The first Modern Humans in Southwestern Iberia and the extinction of Neanderthals. PI: Nuno Bicho. Funding: Wenner-Gren Foundation", "2011-12",
  "The last hunter-gatherers of Muge (Portugal): the origins of social complexity. PI: Nuno Bicho. Funded by the Fundação para a Ciência e Tecnologia - PTDC/HIS‐ARQ/112156/2009", "2012-14"
) %>%
  brief_entries(project, year)


```

# Conference participation

## Oral communications

```{r oral}

bibliography_entries("CV.bib") %>% 
  filter(type == "speech" & `title-short` == "Talk") %>% 
  arrange(!is.na(issued), desc(issued))

```

## Posters

```{r posters}

bibliography_entries("CV.bib") %>% 
  filter(type == "speech" & `title-short` == "Poster") %>% 
  arrange(!is.na(issued), desc(issued))

```

# Fieldwork

## As coordinator

```{r fieldcoord}

tribble(
  ~ project, ~ year,
  "Excavations at the Mesolithic site of Cabeço dos Ossos, Portugal", "2021",
  "Excavations at the Paleolithic site of Gruta do Escoural, Portugal", "2020-",
  "Excavations at the Paleolithic site of Gruta da Companheira, Portugal", "2017-",
  "Excavations at the Paleolithic site of Vale Boi, Portugal", "2009-2019",
  "Excavations at the Mesolithic site of Cabeço da Amoreira, Portugal", "2009-",
  "Excavations at the Mesolithic site of Cabeço da Arruda, Portugal", "2013",
  "Excavations at the Paleolithic sites of Gruta Nova da Columbeira and Lapa do Suão, Portugal", "2009"
) %>%
  brief_entries(project, year)

```

## As invited specialist

```{r fieldinvit}

tribble(
  ~ project, ~ year,
  "DISPERSALS - Resilience, and innovation in Late Pleistocene SE Africa. Lithics specialist. PI: Nuno Bicho", "2023-",
  "Survey and excavations in Mozambique under the project Stone Age Vilankulos: Modern Human Origins Research south of the Rio Save, Mozambique. Lithics specialist and field manager. PI: Jonathan Haws", "2016-19",
  "Excavations at Dehesilla Cave, Spain. Total station specialist. PI: Daniel Garcia Rivero", "2016-19",
  "Survey and excavations in Mozambique under the project Middle Stone Age archaeology and the origins of modern humans in southern Mozambique. Lithics specialist and field manager. PI: Nuno Bicho", "2013-15",
  "Excavations at Picareiro Cave. Total station specialist. PI: Jonathan Haws", "2014-15"
) %>%
  brief_entries(project, year)

```

## Other fieldwork

```{r fieldother}

tribble(
  ~ project, ~ year,
  "Excavations at the Mesolithic site of Cabeço da Amoreira, Portugal. Assistant field manager. PI: Nuno Bicho.", "2008",
  "Testing at the site of Zavial Rockshelter, Portugal. Excavator. PI: Nuno Bicho.", "2006",
  "Excavations at the Paleolithic site of Vale Boi, Portugal. Assistant field manager. PI: Nuno Bicho", "2005-06",
  "Excavations at the Mesolithic site of Barranco das Quebradas, Portugal. Excavator. PI: Maria J. Valente", "2004"
) %>%
  brief_entries(project, year)

```

# Teaching

```{r teaching}

tribble(
 ~ role, ~ company, ~ dates, ~ details, ~ topic,
"Invited Professor", "University of Coimbra, Portugal", "2018-2019", "Field Archaeology (MA in Human Evolution)", "Courses taught:",
"Invited Professor", "Universidade do Algarve, Portugal", "2014-Present", "Introduction to Archaeology (Undergrad. in Archaeology)", "Courses taught:", 
"Invited Professor", "Universidade do Algarve, Portugal", "2014-Present", "Archaeological Theory (MA in Archaeology)", "Courses taught:",
 "Invited Professor", "Universidade do Algarve, Portugal", "2014-Present", "Lithic analysis (MA in Archaeology)", "Courses taught:",
 "Invited Professor", "Universidade do Algarve, Portugal", "2014-Present", "Introduction to lithic analysis (Undergrad. in Archaeology)", "Courses taught:",
 "Invited Professor", "Universidade do Algarve, Portugal", "2014-Present", "Introduction to Data Analysis in Archaeology (Undergrad. in Archaeology)", "Courses taught:",
  "Invited Professor", "Universidade do Algarve, Portugal", "2014-Present", "Prehistoric adaptations to climate change (Ph.D. in Archaeology)", "Courses taught:",
"Invited Professor", "Universidade do Algarve, Portugal", "2014-Present", "Digging data: quantitative and reproducible analysis in Archaeology (Ph.D. in Archaeology)", "Courses taught:",
) %>%
  detailed_entries(with = role, when = dates, where = company, why = details, what = topic)

```

# Administrative and management experience

## Scientific and University management

```{r management}

tribble(
 ~ role, ~ company, ~ dates, ~ topic,
  "Director", "ICArEHB - Universidade do Algarve, Portugal", "2023-", "Interdisciplinary Center for Archaeology and the Evolution of Human Behaviour") %>%
  detailed_entries(with = role, when = dates, where = company, what = topic)

tribble(
 ~ role, ~ company, ~ dates, ~ topic,
  "Invited member", "Universidade do Algarve, Portugal", "2023-", "Quality Assurance Committee") %>%
  detailed_entries(with = role, when = dates, where = company, what = topic)

tribble(
 ~ role, ~ company, ~ dates, ~ topic,
  "Member", "Universidade do Algarve, Portugal", "2023-", "Faculdade de Ciências Humanas e Sociais Scientific Council") %>%
  detailed_entries(with = role, when = dates, where = company, what = topic)

tribble(
 ~ role, ~ company, ~ dates, ~ topic,
  "Member", "Universidade do Algarve, Portugal", "2023-", "Academic Senate") %>%
  detailed_entries(with = role, when = dates, where = company, what = topic)

tribble(
 ~ role, ~ company, ~ dates, ~ topic,
  "Member", "Universidade do Algarve, Portugal", "2023-", "Research Council of the Unidade de Apoio à Investigação Científica e Formação Pós-Graduada") %>%
  detailed_entries(with = role, when = dates, where = company, what = topic)

tribble(
 ~ role, ~ company, ~ dates, ~ topic,
  "Coordinator", "ICArEHB - Universidade do Algarve, Portugal", "2022", "Research Line on Dynamics of Paleolithic people in Europe") %>%
  detailed_entries(with = role, when = dates, where = company, what = topic)

tribble(
 ~ role, ~ company, ~ dates, ~ topic,
  "Coordinator", "ICArEHB - Universidade do Algarve, Portugal", "2019-2021", "Research Line on African Archaeology and Human Evolution") %>%
  detailed_entries(with = role, when = dates, where = company, what = topic)

tribble(
 ~ role, ~ company, ~ dates, ~ topic,
  "Elected member", "Universidade do Algarve, Portugal", "2021-2024", "UAlg's General Council") %>%
  detailed_entries(with = role, when = dates, where = company, what = topic)

```

## Search committees

```{r committees}

tribble(
  ~ committee, ~ year,
  "Call for 1 PhD Fellowship (BI) in Archaeology, expertise in climate modeling, within FINISTERRA: Population trajectories and cultural dynamics of Late Neanderthals in Far Western Eurasia (101045506-FINISTERRA-ERC-2021-COG). Notice nº 046/2023", "2023",
  "Call for 1 PhD Fellowship (BI) in Archaeology, expertise in geoarchaeology, within FINISTERRA: Population trajectories and cultural dynamics of Late Neanderthals in Far Western Eurasia (101045506-FINISTERRA-ERC-2021-COG). Notice nº 037/2023", "2023",
  "Call for 1 MA Fellowship (BI) in Archaeology at ICArEHB. Notice nº 001/2023", "2023",
  "Call for 3 PhD Fellowships (BI) in Archaeology, expertise in isotopes, phytoliths, computational modeling, within DISPERSALS: Dispersals, resilience, and innovation in Late Pleistocene SE Africa, ERC-2021-ADG. Notice nº 039/2023", "2023",
  "Call for 1 Principal Researcher in Archaeology within DISPERSALS: Dispersals, resilience, and innovation in Late Pleistocene SE Africa, ERC-2021-ADG. Notice nº 15831/2023", "2023",
  "Call for 1 Post-Doc Fellowship (BI) in Archaeology, within DIASPORA: Early Human migrations and the Nile Valley: the Kerma region during the MSA. Notice nº 071/2022", "2023",
  "Call for 4 Auxiliar Researchers in Archaeology, expertises in isotopes, phytoliths, computational modeling, biomarkers, within DISPERSALS: Dispersals, resilience, and innovation in Late Pleistocene SE Africa, ERC-2021-ADG. Notice n.º 17876/2022", "2022",
  "Call for 1 post-doctoral fellowship in Archaeology, expertise in fauna taphonomy, within FINISTERRA: Population trajectories and cultural dynamics of Late Neanderthals in Far Western Eurasia (101045506-FINISTERRA-ERC-2021-COG). Notice nº 070/2022", "2022",
  "Call for 1 post-doctoral Assistant Researcher contract in Archaeology at ICArEHB, through CEEC institutional program. Notice nº 202205/0517", "2022",
  "Call for 1 PhD Fellowship (BI) in Archaeology, expertise in lithic analysis, within FINISTERRA: Population trajectories and cultural dynamics of Late Neanderthals in Far Western Eurasia (101045506-FINISTERRA-ERC-2021-COG). Notice nº 034/2022", "2022",
  "Call for 1 PhD Fellowship (BI) in Archaeology, expertise in microfauna analysis, within FINISTERRA: Population trajectories and cultural dynamics of Late Neanderthals in Far Western Eurasia (101045506-FINISTERRA-ERC-2021-COG). Notice nº 035/2022", "2022",
  "Call for 1 post-doctoral fellowship in Archaeology, expertise in isotope analysis, within FINISTERRA: Population trajectories and cultural dynamics of Late Neanderthals in Far Western Eurasia (101045506-FINISTERRA-ERC-2021-COG). Notice nº 039/2022", "2022",
  "Call for 1 post-doctoral Assistant Researcher contract in Archaeology at ICArEHB, through CEEC institutional program", "2022",
  "Call for 1 MA fellowship in Archaeology within the project ALG-45-2020-41– SciTour - Turismo Científico: uma nova abordagem para promover o turismo no Algarve", "2022",
  "Call for 1 Research Fellowship (BI) in Archaeology within the project DIASPORA – As primeiras migrações humanas no vale do Nilo: a região de Kerma durante o MSA (PTDC/HAR-ARQ/0131/2020). Notice nº 024/2022", "2022",
  "Call for 1 Research Fellowship (BI) in Archaeology at ICArEHB (UID/ARQ/04211/2019). Notice nº 057/2021", "2021",
  "Call for 1 Research Fellowship (BI) in Archaeology within The Muge Shellmiddens Project: A new Portal to the last hunter-gatherers the Tagus Valley, Portugal (ALG-01-0145-FEDER-29680). Notice nº 033/2021.", "2021",
  "Call for 4 Ph.D. fellowships (BD) in Prehistoric Archaeology at ICArEHB (UID/ARQ/04211/2019)", "2021",
  "Call for 1 post-doctoral contract in Archaeology at ICArEHB (UID/ARQ/04211/2019)", "2020",
  "Call for 1 Research Fellowship (BI) in 3D Animation within The Muge Shellmiddens Project: A new Portal to the last hunter-gatherers the Tagus Valley, Portugal (ALG-01-0145-FEDER-29680). Notice nº 079/2019", "2020",
  "Call for 1 Research Fellowship (BI) in Computer Science and engineering within The Muge Shellmiddens Project: A new Portal to the last hunter-gatherers the Tagus Valley, Portugal (ALG-01-0145-FEDER-29680). Notice nº 001/2020", "2020",
  "Call for 1 Research Fellowship (BI) in 3D Animation within The Muge Shellmiddens Project: A new Portal to the last hunter-gatherers the Tagus Valley, Portugal (ALG-01-0145-FEDER-29680). Notice nº 002/2020", "2020",
  "Call for 1 Research Fellowship (BI) in Design within The Muge Shellmiddens Project: A new Portal to the last hunter-gatherers the Tagus Valley, Portugal (ALG-01-0145-FEDER-29680). Notice nº 018/2020", "2020",
  "Call for 1 Research Fellowship (BI) in Archaeology within The origins and evolution of human cognition and the impact of Southwestern European coastal ecology (ALG-01-0145-FEDER-27833). Notice nº 060/2020", "2020",
  "Call for 1 funding officer position at ICArEHB (UID/ARQ/04211/2019)", "2020",
  "Call for 1 administrative position at ICArEHB (UID/ARQ/04211/2019). Notice nº 046/2019", "2019",
  "Call for 1 post-doctoral contract within the project FLAME – Fire among Anatomically Modern Humans and Neanderthals as revealed from Archaeological Microstratigraphic evidence", "2019",
  "Call for 1 post-doctoral contract within the project The origins and evolution of human cognition and the impact of Southwestern European coastal ecology (ALG-01-0145-FEDER-27833)", "2019"
) %>%
  brief_entries(committee, year)

```

# Extension and knowledge dissemination

## Student and Postdoctoral supervision

```{r supervision}

tribble(
  ~ student, ~ year,
  "Nolan Ferar (Ph.D. dissertation, ICArEHB, Universidade do Algarve, Portugal) - Funded by FCT", "2022-",
  "Mpumi Maringa (Ph.D. dissertation, ICArEHB, Universidade do Algarve, Portugal", "2022-",
  "Jovan Galfi (Ph.D. dissertation, ICArEHB, Universidade do Algarve, Portugal", "2020-",
  "Joana Belmiro (Ph.D. dissertation, Universidade do Algarve, Portugal) - Funded by FCT", "2020-",
  "Daniela Maio (Ph.D. dissertation, Universidade do Algarve, Portugal)", "2019-",
  "Emily Hallinan (Marie Curie Fellow, ICArEHB, Universidade do Algarve", "2020-2022",
  "Pedro Horta (Ph.D. dissertation, Universidade do Algarve, Portugal) - Funded by FCT", "2017-2022",
  "Joana Belmiro (M.A. dissertation, Universidade do Algarve, Portugal)", "2017-2019",
  "Joana Belmiro (Undergrad. dissertation, Universidade do Algarve, Portugal)", "2017",
  "Pedro Horta (M.A. dissertation, Universidade do Algarve, Portugal)", "2014-15"
) %>%
  brief_entries(student, year)

```

## Invited seminars

```{r seminars}

tribble(
  ~ talk, ~ year,
  '"The Upper Paleolithic occupations at Vale Boi (Vila do Bispo, Algarve)". Associação Arqueológica do Algarve, São Brás de Alportel/Lagoa', "2024",
  '"40 mil anos depois: o que sabemos sobre o desaparecimento dos Neandertais". Academia das Ciências de Lisboa, Lisbon', "2023",
  '"O que sabemos sobre o desaparecimento dos Neandertais?". FI.CA, Lisbon', "2022",
  '"Human adaptations to the Last Glacial Maximum in the Iberian Peninsula". Hebrew University MA seminar series "The social dimensions of human adaptation to extreme environments", Online', "2021",
  '"Proto-Solutrean and Solutrean adaptations in westernmost Iberia: the Vale Boi and Lapa do Picareiro case-studies". Erlangen, DE (with Nuno Bicho, Joana Belmiro and Jonathan Haws)', "2019",
  '"A última decada de trabalhos arqueológicos no concheiro do Cabeço da Amoreira, Muge". Muge, Salvaterra de Magos, PT (with Célia Gonçalves, Nuno Bicho and Lino André)', "2019",
  '"Evolução humana e a tecnologia dos instrumentos em pedra". Universidade de Coimbra, Coimbra, Portugal, Universidade de Coimbra, Coimbra, PT.', "2017",
  '"Neanderthal to Modern Humans: the transition in Portugal". Algarve Archaeological Association, São Brás de Alportel, PT', "2017",
  '"Prehistoric Hunter-Gatherer Adaptations in Southwestern Iberia and Mozambique". Max Planck Institute for the Science of Human History, Jena, DE', "2016",
  '"The end of the Middle Paleolithic and the arrival of anatomically modern humans to Southwestern Iberia". Max Planck Institute, Leipzig, DE (with Nuno Bicho)', "2016",
  '"Recent developments in the analysis of lithic technology". University of Seville, ES', "2016",
  '"Laboratorial techniques for the analysis of lithic industries". University Eduardo Mondlane, Maputo, MZ', "2015",
  '"Abrupt climate change and human adaptations". University Eduardo Mondlane, Maputo, MZ', "2014",
  '"Stone Age archaeology in the Niassa region, Mozambique. Preliminary results from the 2014 campaign". Geographical Society of Lisbon, Lisbon, PT (with N. Bicho, J. Haws, O. Madime, C. Gonçalves, V. Aldeias, and M. Benedetti)', "2014",
  '"Human responses to environmental change: the case of the Portuguese Upper Paleolithic". National Museum of Archaeology, Lisbon, PT', "2014",
  '"The mediterranean influence on the social networks of the Solutrean in Iberia". National Museum of Archaeology, Lisbon, PT', "2014"
) %>%
  brief_entries(talk, year)

```

## Conferences/Symposia Organized

```{r symposia}
tribble(
  ~ talk, ~ year,
  '"Symposium Expedient Technological Behavior: Global Perspectives and Future Directions". With Nolan Ferar. SAA 2024, New Orleans, USA', "2024",
  '"European Society for Human Evolution Annual Meeting". Universidade do Algarve, Faro, PT', "2018",
  '"International Congress The Solutrean". With Isabell Schmidt, Nuno Bicho and Gerd-Christian Weniger. Universidade do Algarve, Faro, PT', "2017",
  '"Symposium Global perspectives on the impact of abrupt climate change in hunter-gatherer-technologies". With Nuno Bicho. SAA 2016, Orlando, USA', "2016",
  'Conference cycle "Arqueologia às 6". Organized by ICArEHB. Faro, PT', "2015-2017",
  '"Symposium Solutrean points: past and present". With Juan Gibaja Bao. At the International Conference El Solutrense. Almería, ES', "2012",
  '"IV International Conference of Young Researchers in Archaeology (JIA 2011)". Universidade do Algarve, Faro, PT', "2011",
  '"Session Demographic processes and cultural change: archaeological perspectives". III International Conference of Young Researchers in Archaeology (JIA 2010). Barcelona, ES', "2010"
) %>%
  brief_entries(talk, year)

```

## Workshops and Exhibitions Organized

```{r workshops}
tribble(
  ~ talk, ~ year,
  'Workshop "Data Analysis and R Statistics". Universidade do Algarve, Faro, PT', "2016",
  'Exhibition "The Prehistoric Origins of the Algarve". Organized by ICArEHB and Centro de Ciência Viva do Algarve. Faro, Tavira, Lagos and Vila do Bispo. PT', "2016",
  'Exhibition "At the End of the World 30 thousand years ago: the prehistoric hunter-gatherers of Vale Boi". Organized by ICArEHB. Vila do Bispo, PT', "2014",
  'Workshop "Total Station in Archaeology". Universidade do Algarve, PT. Promoted by NAP', "2011",
  'Exhibition "On the margins of the past: the Muge shell middens". Salvaterra de Magos, PT. Organized by NAP', "2011",
  'Conferences "Arqueologia ao Sul". Universidade do Algarve, Faro, PT. Organized by NAP', "2010-14",
  'Workshop "Analysis of Lithic Industries" (Universidade do Algarve, Faro, PT. Promoted by NAP', "2010",
  'Workshop "NESPOS On Tour". Universidade do Algarve, Faro PT. Promoted by NAP', "2010"
) %>%
  brief_entries(talk, year)
```

## Tutoring

```{r tutoring}
tribble(
  ~ talk, ~ year,
  'Instructor of Statistics in Archaeology using R in one the Training in Frontier Archaeology ICArEHB actions', "2022",
  'Creation of the archaeological survey and excavation Massive Open Online Courses (MOOC) modules of the *Online Learning on African Archeology and Heritage (ONLAAH)* project', "2019-") %>%
  brief_entries(talk, year)
```

# Additional Training

-   Workshop Introduction to Open Science. Universidade do Algarve, Faro, PT.
-   Workshop Machine Learning in Archaeology. Universidade do Algarve, Faro, PT:
-   Workshop Science and public policy: how to build bridges?. Universidade do Algarve, Faro, PT.
-   Workshop Data Analysis and R Statistics. Universidade do Algarve, Faro, PT.
-   Workshop EVAN Toolbox. Max Planck Institute for Evolutionary Abthropology, Leipzig, DE.
-   Workshop NESPOS on Tour. Universidade do Algarve, Faro, PT.
-   Workshop Understanding tool use, multidisciplinary perspectives on the cognition and ecology of tool using behaviors. Max Planck Institute for Evolutionary Anthropology, Leipzig, DE.
-   Workshop Tools in organic materials: methods of analysis. CSIC, Barcelona, ES.
-   Workshop Advanced SPSS applications. Universidade do Algarve, Faro, PT.
-   Workshop Functional studies in non-siliceous rocks. Lisbon, PT.

# Reviewer for academic journals and funding agencies

-   Fund for Scientific Research -- FNRS Belgium (1)
-   Nature Scientific Reports (1)
-   PLoS ONE (4)
-   Journal of Human Evolution (2)
-   Advances in Archaeological Practice (2)
-   Journal of Archaeological Science (1)
-   Journal of Archaeological Science: Reports (2)
-   Quaternary International (1)
-   Journal of Paleolithic Archaeology (3)
-   African Archaeological Review (1)
-   Antiquity (1)
-   National Science Centre Poland (1)
-   Paleoanthropology Journal (1)
-   Journal of Open Archaeology Data (1)
-   Journal of Environmental Archaeology (1)

# Professional memberships

-   European Society for the Study of Human Evolution
-   Society for American Archaeology
-   Paleoanthropology Society
-   Archaeological Institute of America
-   Seminário Permanente dos Jovens Cientistas do Instituto de Altos Estudos da Academia das Ciências de Lisboa, Membro.

# Languages

-   Portuguese (native)
-   English (proficient)
-   Spanish (conversational)
